\ttitle

### 2015-08-05 Chapters/Introduction.tex
Introduction - Spelling and grammar

Aside from a thorough history, clinical examination, blood tests including liver function tests, diagnosis requires cross-sectional imaging in the form of a contrast-enhanced computerised tomogram (CECT) of the abdomen using a pancreas-specific protocol (a modified form of the portal-venous phase). CECT of the pancreas when combined with CT Thorax also provides accurate information on staging of the disease with regards to metastasis and this can be supplemented by further imaging such as Positron Emission Tomography (PET-CT) or contrast-enhanced MRI Liver in specific cases. CECT-pancreas is also useful for assessing local resectability with regards to vascular involvement. Endoscopic ultrasound (EUS) is also useful in assessing vascular involvement and for obtaining tissue samples for histological examination. In jaundiced patients, endoscopic retrograde cholangio pancreatography (ERCP) plays an important role in the alleviation of jaundice by placing stents across the obstructed bile ducts, accurate visualisation of the biliary anatomy as well as obtaining brushings from within the bile ducts for cytological examination. The role of preoperative biliary drainage is discussed in more detail in section %\ref{sec:preoperative_biliary_drainage].

@@

Reconstruction was performed as follows: Either a pancreatico-jejunostomy was performed using 4-0 Biosyn sutures in a two-layer duct-to-mucosa technique or a pancreatico-gastrostomy was performed using 3/0 Biosyn sutures placed in a similar manner. Hepaticojejunostomy was performed using interrupted 4/0 Biosyn sutures while the gastrojeunonostomy or duodenojejunostomy (in PPPD) was performed using continuous 3/0 PDS sutures in a 2-layers. One or two surgical drains were placed and the abdomen was closed after ensuring haemostasis.

\subsection{Postoperative care}
All patients were routinely admitted to the Surgical High Dependency Unit unless intra-operative events necessitated admission to the Intensive Care Unit. A standardised regimen of intravenous fluids, naso-jejunal feeding, mobilisation and physiotherapy was implemented in all patients. Standard physiological parameters including haemodynamic parameters, renal function and arterial blood gases were used to monitor adequate end organ perfusion. All patients received proton pump inhibitors and octreotide. Patients were discharged to the general surgical ward as early as possible.

@@

Postoperative pancreatic fistula is one of the most dreaded complications after a pancreaticoduodenectomy and can be associated with significant short-term morbidity as well as long-term disability. The reported incidence of postoperative pancreatic fistula varies from 2\% to 30\% after pancreaticoduodenectomy.\parencite{yeo_six_1997,deoliveira_assessment_2006,bassi_postoperative_2005,winter_biochemical_2007,pratt_risk_2008} 
The variation in reported incidence has been largely due to lack of clear definition of what constituted a postoperative pancreatic fistula. It can be a result of breakdown or poor healing at the pancreaticojejunostomy/pancreaticogastrostomy or may be the result of direct parenchymal leak unrelated to the anastomosis. It is now generally accepted that 1 in 4 patients will develop a pancreatic fistula as defined by the International Study Group for Pancreatic Fistula (ISGPF) which has published a consensus statement on the definition and grading of postoperative pancreatic fistula.\parencite{bassi_postoperative_2005} A postoperative pancreatic fistula is defined as drain output of any measurable quantity after the third postoperative day with amylase content greater than three times the upper limit of the normal serum amylase value at the laboratory used for testing. Three grades of postoperative pancreatic fistula have been defined based on clinical severity as described in Table \ref{table:isgps_popf} on p\pageref{table:isgps_popf}. Grade B and C fistulae are considered to be clinically significant as they alter patient management and are often associated with other secondary complications such as intra-abdominal sepsis, post-pancreatectomy haemorrhage or delayed gastric emptying. They may require further intervention (either radiological or operative) and may result in prolonged critical care support.

@@

Haemorrhage is often secondary to non-healing of the pancreatico-jejunal anastomosis leading to leakage of amylase-rich pancreatic juices into the retroperitoneum or secondary to intra-abdominal sepsis or bile leak.\parencite{tien_risk_2005, koukoutsis_haemorrhage_2006, choi_delayed_2004, balladur_bleeding_1996} This can then lead to erosion of ligated blood vessels, most commonly the stump of the gastro-duodnenal artery. Post-pancreatectomy haemorrhage is often managed with angiographic embolisation of the bleeding vessel and surgical intervention is only rarely required. The grading of severity of post-pancreatectomy haemorrhage as described by the International Study Group of Pancreatic Surgery\parencite{wente_postpancreatectomy_2007} is shown in Table \ref{table:isgps_pph} on p\pageref{table:isgps_pph}.

@@

The ESPAC-1 trial compared the use of chemotherapy with fluorouracil and folinic acid or chemoradiotherapy with 20Gy radiotherapy and 5-fluorouracil (5-FU) versus observation only in patients who underwent potentially curative surgery for pancreatic ductal adenocarcinoma. The 5-year survival rate in patients who received chemotherapy was 20\% while it was 8\% in patients who did not receive demonstrating a clear survival benefit for adjuvant chemotherapy. However, adjuvant chemoradiotherapy resulted in 5-year survival dropping from 20\% to 10\% and has largely been abandoned in the adjuvant setting. The ESPAC-3 trial compared the efficacy of 5-FU and folinic acid with gemcitabine and reported that gemcitabine was not superior to 5-FU/folinic acid.

@@

\subsection{Risk Stratification}
\todo{Intro - risk stratification}
%Include other measures - discuss POSSUM, APACHE, mGPS, EPASS and everything else here.
Physiological fitness or reserve may be defined as the ability of the patient's organ systems to respond appropriately and adequately to the stress of surgery. Major surgery has a significant physiological impact on multiple organ systems, especially the cardiorespiratory system. The ability of the cardiorespiratory system as well as other physiological systems including renal, gastrointestinal, hepatic, coagulatory and immunological systems to cope with major surgery and its sequelae plays a major role in determining postoperative outcomes. Therefore, accurate assessment of physiololgical fitness is important in selecting patients who will benefit from surgery, advising patients of the risks of surgery as it applies to them, preoperative optimisation as well as in perioperative management decisions. A thorough clinical history of co-existing illnesses, general examination and review of patient's medications is followed by more specialist testing to assess fitness. Numerous methods including individual tests and composite scores derived from physiological and biochemical variables have been used for risk assessment.

@@

\subsection{Origins and physiological basis of cardiopulmonary exercise testing}
The measurement of aerobic fitness has been an important aspect of improving atheletic performance in competitive sports. Cardiopulmonary exercise testing has been used for a long time for detailed measurement of aerobic fitness in athletes as well as offer customised training regimens aimed at improving performance at activities that place a signficant stress on the cardiovascular and respiratory systems.

@@

Spirometry was performed in all patients prior to CPET. Capillary blood gases were measured in all patients after CPET. An electronically braked cycle ergometer was used to increase resistance to pedalling in preset increments. A tight-fitting face mask was placed on the patient covering the nose and the mouth. This allowed breath-by-breath gas analysis thus allowing measurement of several respiratory parameters as listed in table []. 12-lead electocardiogram was recorded at the same time.

The test started with an initial 3-minute rest period to allow measurement of baseline parameters. This was followed by an incremental work-load test that involved the patient pedalling approximately at 60 revoultions per minute while the resistance to pedalling was gradually increased in preset increments. The test was terminated when patients reached volitional fatigue (maximal exercise tolerance), significant ischaemic changes on ECG or for other safety reasons.

@@

\subsection{End-tidal $O_2$ and $CO_2$, $P_{ET_{O_2}}$, $P{ET_{CO_2}}$}
$PET_{O_2}$ and $PET_{CO_2}$ are the partial pressures of $O_2$ and $CO_2$ at the end of an exhaled breath and are closely related to PaO2 and PaCO2 respectively. $P{ET_{CO_2}}$ is dependent on pulmonary gas-exchange which is in turn influenced by the right ventricular output, pulmonary blood flow and alveolar gas exchange. The changes in $P_{ET_{O_2}}$ and $P{ET_{CO_2}}$ during exercise help identify ventilation-perfusion mismatch as well as hyperventilation.

@@

\subsection{Breathing frequency, $B_f$}
In healthy individuals, breathing frequency is one of the most important respiratory responses to exercise in order to increase oxygen uptake. However, a variety of respiratory disorders including obstructive disorders such as chronic obstructive pulmonary disease or restrictive disorders such as interstitial lung disease can impair this response. Morbidly obese patients may also have additional restrictions on respiratory response due to their body habitus. On the other hand, hyperventilation in an otherwise healthy patient may make interpretation of the results difficult although the analysis of trends in other parameters such as $P_{ET_{O_2}}$ and $P{ET_{CO_2}}$ aids in the diagnosis of hyperventilation versus true respiratory problems.

@@

These findings have since been replicated in several other studies although the threshold value of $\dot{V}_{O_2}$AT as well as the inclusion of other CPET parameters to attribute risk differ in these studies. Hightower and co-workers studied 32 patients over the age of 18 undergoing a variety of major elective abdominal surgery. In this heterogenous cohort, they reported that the percent predicted anaerobic threshold achieved $<$ 75\%, heart rate at the anaerobic threshold and the heart rate response from rest to anaerobic thresold were all independently associated with postoperative morbidity. While it is difficult to extrapolate results from this small heterogenous cohort to the general surgical population, it would appear that other cardiopulmonary exercise test parameters may have a role in predicting risk in these patients.\parencite{hightower_pilot_2010}

Snowden and coworkers reported in a study of 171 patients of who 123 underwent surgery, that $\dot{V}_{O_2}AT<10.1$ ml/kg/min was associated with not only cardiovascular complications but also with other complications including pulmonary, renal, gastrointestinal, infective and haematological complications. One of the strengths of this study was the fact that the clinicians treating these patients were blinded to the results of preoperative cardiopulmonary exercise testing results thus avoiding management bias. They also included other cardiopulmonary exercise test parameters such as $\dot{V}_{O_2}$Peak and $\dot{V}_E/\dot{V}_{CO_2}$ at the anaerobic threshold in addition to $\dot{V}_{O_2}$AT in their analysis and showed that $\dot{V}_{O_2}$AT was more predictive of postoperative morbidity than POSSUM derived morbidity, cardiac risk scoring index or a validated activity questionnaire. However, this study also included a heterogenous cohort of diseases including vascular, pancreatic and hepatobiliary disorders and sarcomas.\parencite{snowden_submaximal_2010}

@@

\subsection{Systemic inflammation and long-term survival}
Systemic inflammation is associated with poorer survival in patients undergoing potentially curative surgery for pancreatic cancer \parencite{jamieson_systemic_2005,clark_preoperative_2007,bhatti_preoperative_2010} as well as in patients with inoperable pancreatic cancer.\parencite{glen_evaluation_2006} Patients with ductal adenocarcinoma of the head of the pancreas undergoing potentially curative resection survived for a median of 21.5 months if their CRP was $\leq$ 10 mg/dl a month after their surgery but only 8.4 months if their CRP remained persistently elevated at over 10 mg/dl approximately a month after their operation.\parencite{jamieson_systemic_2005} Similar findings have been reported in cancers involving other organs using both the mGPS and other scores such as the neutrophil-lymphocyte ratio (NLR). A selection of these studies are presented in Table %Create new table here for this.

@@

Preoperative systemic inflammation has been reported to be associated with infectious complications in patients undergoing potentially curative surgery for colorectal cancer.\parencite{moyes_preoperative_2009} In a study of 455 patients, Moyes and coworkers reported that an elevated preoperative modified Glasgow Prognostic Score (\ref{table:mGPS}) was associated with increased incidence of infectious complications in patients undergoing elective as well emergency colorectal cancer surgery. They postulated that several mechanisms may have a role including disregulation of cell-mediated immunity, impaired T-lymphocyte response, disorders in the complement pathway and possibly due to loss of lean tissue and protein as a consequence of systemic inflammation. Preoperative mGPS has also been shown to predict postoperative morbidity in patients undergoing oesophageal resection for cancer.\parencite{vashist_glasgow_2010}

@@

In a study of 383 patients undergoing elective rectal cancer surgery with primary anastomosis, Welsch and co-workers reported that persistently raised CRP level over 140 mg/L after the third/fourth postoperative day was associated with anastomotic leak.\parencite{welsch_c-reactive_2007} They also reported in a separate study of 688 patients undergoing pancreatic resection with pancreaticojejunostomy for neoplastic disease or chronic pancreatitis, that persistently elevated CRP levels greater than 140 mg/L on the fourth postoperative day was associated with increased incidence of complications.

@@

While CRP level between the third and fifth postoperative day has been reported to be most predictive of complications,  the complications themselves do not become clinically apparent until a later in the postperative course, ofter after the fifth postoperative period. This has led some authors to postulate that the elevated CRP levels may in fact be due to an abnormally modulated postoperative inflammatory response resulting in an initial exaggerated systemic inflammatory response syndrome (SIRS) followed by a compensatory anti-inflammatory response syndrome (CARS).

@@

This form of 'immunoparalysis' was first described in patients after major trauma with tissue damage \parencite{abraham_effects_1985,bandyopadhyay_negative_2007} or after haemorrhage on its own without associated tissue trauma. \parencite{stephan_hemorrhage_1987} In a detailed review of the mechanisms underlying the compensatory anti-inflammatory response syndrome, Ward and coworkers describe SIRS and CARS to be mirror images suggesting that a disproportionately high SIRS is followed by a period of immunosuppression that leaves the patient prone to further complications.\parencite{ward_compensatory_2008}

@@

Yamaguchi and co-workers compared the levels of pro- and anti-inflammatory cytokines in patients undergoing cholecystectomy versus patients undergoing trans-thoracic oesophagectomy. They reported that the initial inflammatory phase was followed by an immunosuppressive phase that started around the seventh postoperative day in patients undergoing oesophagectomy. However, patient who underwent underwent an open cholecystectomy did not experience this immunosuppressive phase, leading them to postulate that the degree of immunosuppression was directly proportional to the intial pro-inflammatory process. This in turn was related to the greater degree of surgical stress and tissue trauma that occurs with a trans-thoracic oesophagectomy. They also reported that in a randomised cohort that received an infusion of lymphokine-activated natural killer cells immediately after oesophagectomy, there was a trend towards fewer infectious complications.\parencite{yamaguchi_postoperative_2006}

@@

The detrimental effects of jaundice on the heart and the circuilatory system has been recognised for over 150 years now. King and coworkers reported that the intravenous injection of bile caused bradycardia, hypotension and eventually death in dogs.\parencite{king_effect_1909} The concept of a \textit{'jaundiced heart'} was fist put forth by Green and coworkers in 1986.\parencite{green_jaundiced_1986} They performed choledocho-caval anastomoses in 5 dogs and studied cardiac function before and 2 weeks after this procedure. They reported that 'cholemia' was associated with impaired left ventricular function and blunted reponse to sympathomimmetic agents. Similar findings have been reported by other authors in animal studies. \parencite{binah_obstructive_1985,bomzon_systemic_1986}

@@

\subsection{Role of preoperative biliary drainage}
%fang_pre-operative_2012 - Cochrane review
%wang_preoperative_2008 - Cochrane review
The increased incidence of adverse events after surgery in patients with obstructive jaundice resulted in routine preoperative biliary drainage becoming the standard practice in patients undergoing major pancreatic surgery. Moreover, while significant improvements in perioperative care has resulted in reduction in postoperative mortality, morbidity rates have remained high. Obstructive jaundice has provided a potential target for preoperative optimisation in an attempt to improve postoperative outcomes.

Early observational studies of small cohorts of patients challenged this paradigm reporting that surgery in jaundiced patients was not associated ith increased risk and in fact preoperative biliary drainage increased hospitalisation costs. \parencite{snellen_influence_1985, bakkevold_morbidity_1993, pitt_does_1985} Hatfield and co-workers reported in one of the earliest randomised controlled trial comparing pancreaticoduodenectomy with or without preoperative biliary drainage in 57 patients that there was no difference in postoperative complications or mortality but increased biliary drainage associated complications. \parencite{hatfield_preoperative_1982}

@@

A recent randomised controlled trial of 202 patients reported that there was no significant difference in post-operative morbidity (47\% vs, 37\%, p=0.14), mortality or length of stay between the group that underwent preoperative biliary drainage and the group that underwent early surgery. There was a high incidence of procedure related complications in the drainage group (46\%). The high incidence of procedure related complications, the use of plastic stents, the 25\% failure to drain at first attempt and exclusion of patients with serum bilirubin $>$ 250 $\mu$mol/L were significant limitations of the study. However, the study does show that early surgery in the jaundiced patient with serum bilirubin $<$ 250 $\mu$mol/L is safe and not associated with increased morbidity or mortality as once feared. Mezhir and co-workers also reported a high (23\%) stent-related complication rate in their retrospective study of 340 patients where 96\% of 201 patients in the biliary drainage group had plastic stents placed. However, they also reported a greater incidence of wound infections (7\% vs 20\%, p=0.01) and intra-abdominal abscesses (3\% vs 12\%, p=0.03) in patients who underwent biliary drainage.

@@

The impact of obesity on the rate of postoperative complications remains controversial. It has been a commonly held belief that obese patients had greater incidence of complications. Some of the adverse outcomes such as wound complications and respiratory dyfunction where conseidered to be the direct result of the increased body mass index while other complications where related to the associated comorbidities such as cardiovascular disease, obstructive sleep apnea or diabetes mellitus.

>>>>

Aside from a thorough history, clinical examination, blood tests including liver function tests, diagnosis requires cross-sectional imaging in the form of a contrast-enhanced computerised tomogram (CECT) of the abdomen using a pancreas-specific protocol (a modified form of the portal-venous phase). CECT of the pancreas when combined with CT Thorax also provides accurate information on staging of the disease with regards to metastasis and this can be supplemented by further imaging such as Positron Emission Tomography (PET-CT) or contrast-enhanced MRI Liver in specific cases. CECT-pancreas is also useful for assessing local resectability with regards to vascular involvement. Endoscopic ultrasound (EUS) is also useful in assessing vascular involvement and for obtaining tissue samples for histological examination. In jaundiced patients, endoscopic retrograde cholangio pancreatography (ERCP) plays an important role in the alleviation of jaundice by placing stents across the obstructed bile ducts, accurate visualisation of the biliary anatomy as well as obtaining brushings from within the bile ducts for cytological examination. The role of preoperative biliary drainage is discussed in more detail in Section \ref{sec:preoperative_biliary_drainage}.

@@

Reconstruction was performed as follows: Either a pancreatico-jejunostomy was performed using 4-0 Biosyn sutures in a two-layer duct-to-mucosa technique or a pancreatico-gastrostomy was performed using 3/0 Biosyn sutures placed in a similar manner. Hepatico-jejunostomy was performed using interrupted 4/0 Biosyn sutures while the gastro-jejunostomy or duodenojejunostomy (in PPPD) was performed using continuous 3/0 Polydioxanone (PDS) sutures in 2-layers. One or two surgical drains were placed and the abdomen was closed after ensuring haemostasis.

\subsection{Postoperative care}
All patients were routinely admitted to the Surgical High Dependency Unit unless intra-operative events necessitated admission to the Intensive Care Unit. A standardised regimen of intravenous fluids, naso-jejunal feeding, mobilisation and physiotherapy was implemented in all patients. Standard physiological parameters including haemodynamic parameters, renal function and arterial blood gases were used to monitor adequate end organ perfusion. All patients received proton pump inhibitors and Octreotide Patients were discharged to the general surgical ward as early as possible.

@@

Postoperative pancreatic fistula is one of the most dreaded complications after a pancreaticoduodenectomy and can be associated with significant short-term morbidity as well as long-term disability. The reported incidence of postoperative pancreatic fistula varies from 2\% to 30\% after pancreaticoduodenectomy.\parencite{yeo_six_1997,deoliveira_assessment_2006,bassi_postoperative_2005,winter_biochemical_2007,pratt_risk_2008} 
The variation in reported incidence has been largely due to lack of clear definition of what constituted a postoperative pancreatic fistula. It can be a result of breakdown or poor healing at the pancreatico-jejunostomy/pancreatico-gastrostomy or may be the result of direct parenchymal leak unrelated to the anastomosis. It is now generally accepted that 1 in 4 patients will develop a pancreatic fistula as defined by the International Study Group for Pancreatic Fistula (ISGPF) which has published a consensus statement on the definition and grading of postoperative pancreatic fistula.\parencite{bassi_postoperative_2005} A postoperative pancreatic fistula is defined as drain output of any measurable quantity after the third postoperative day with amylase content greater than three times the upper limit of the normal serum amylase value at the laboratory used for testing. Three grades of postoperative pancreatic fistula have been defined based on clinical severity as described in Table \ref{table:isgps_popf} on p\pageref{table:isgps_popf}. Grade B and C fistulae are considered to be clinically significant as they alter patient management and are often associated with other secondary complications such as intra-abdominal sepsis, post-pancreatectomy haemorrhage or delayed gastric emptying. They may require further intervention (either radiological or operative) and may result in prolonged critical care support.

@@

Haemorrhage is often secondary to non-healing of the pancreatico-jejunal anastomosis leading to leakage of amylase-rich pancreatic juices into the retroperitoneum or secondary to intra-abdominal sepsis or bile leak.\parencite{tien_risk_2005, koukoutsis_haemorrhage_2006, choi_delayed_2004, balladur_bleeding_1996} This can then lead to erosion of ligated blood vessels, most commonly the stump of the gastro-duodenal artery. Post-pancreatectomy haemorrhage is often managed with angiographic embolisation of the bleeding vessel and surgical intervention is only rarely required. The grading of severity of post-pancreatectomy haemorrhage as described by the International Study Group of Pancreatic Surgery\parencite{wente_postpancreatectomy_2007} is shown in Table \ref{table:isgps_pph} on p\pageref{table:isgps_pph}.

@@

The ESPAC-1 trial compared the use of chemotherapy with Fluorouracil and folinic acid or chemoradiotherapy with 20Gy radiotherapy and 5-Fluorouracil (5-FU) versus observation only in patients who underwent potentially curative surgery for pancreatic ductal adenocarcinoma. The 5-year survival rate in patients who received chemotherapy was 20\% while it was 8\% in patients who did not receive demonstrating a clear survival benefit for adjuvant chemotherapy. However, adjuvant chemoradiotherapy resulted in 5-year survival dropping from 20\% to 10\% and has largely been abandoned in the adjuvant setting. The ESPAC-3 trial compared the efficacy of 5-FU and Folinic acid with Gemcitabine and reported that Gemcitabine was not superior to 5-FU/Folinic acid.

@@

\subsection{Risk Stratification}
\todo{Intro - risk stratification}
%Include other measures - discuss POSSUM, APACHE, mGPS, EPASS and everything else here.
Physiological fitness or reserve may be defined as the ability of the patient's organ systems to respond appropriately and adequately to the stress of surgery. Major surgery has a significant physiological impact on multiple organ systems, especially the cardiorespiratory system. The ability of the cardiorespiratory system as well as other physiological systems including renal, gastrointestinal, hepatic, coagulatory and immunological systems to cope with major surgery and its sequelae plays a major role in determining postoperative outcomes. Therefore, accurate assessment of physiological fitness is important in selecting patients who will benefit from surgery, advising patients of the risks of surgery as it applies to them, preoperative optimisation as well as in perioperative management decisions. A thorough clinical history of co-existing illnesses, general examination and review of patient's medications is followed by more specialist testing to assess fitness. Numerous methods including individual tests and composite scores derived from physiological and biochemical variables have been used for risk assessment.

@@

\subsection{Origins and physiological basis of cardiopulmonary exercise testing}
The measurement of aerobic fitness has been an important aspect of improving athletic performance in competitive sports. Cardiopulmonary exercise testing has been used for a long time for detailed measurement of aerobic fitness in athletes as well as offer customised training regimens aimed at improving performance at activities that place a significant stress on the cardiovascular and respiratory systems.

@@

Spirometry was performed in all patients prior to CPET. Capillary blood gases were measured in all patients after CPET. An electronically braked cycle ergometer was used to increase resistance to pedalling in preset increments. A tight-fitting face mask was placed on the patient covering the nose and the mouth. This allowed breath-by-breath gas analysis thus allowing measurement of several respiratory parameters as listed in table []. 12-lead electrocardiogram was recorded at the same time.

The test started with an initial 3-minute rest period to allow measurement of baseline parameters. This was followed by an incremental work-load test that involved the patient pedalling approximately at 60 revolutions per minute while the resistance to pedalling was gradually increased in preset increments. The test was terminated when patients reached volitional fatigue (maximal exercise tolerance), significant ischaemic changes on ECG or for other safety reasons.

@@

\subsection{End-tidal $O_2$ and $CO_2$, $P_{ET_{O_2}}$, $P{ET_{CO_2}}$}
$PET_{O_2}$ and $PET_{CO_2}$ are the partial pressures of $O_2$ and $CO_2$ at the end of an exhaled breath and are closely related to $P_aO_2$ and $P_aCO_2$ respectively. $P{ET_{CO_2}}$ is dependent on pulmonary gas-exchange which is in turn influenced by the right ventricular output, pulmonary blood flow and alveolar gas exchange. The changes in $P_{ET_{O_2}}$ and $P{ET_{CO_2}}$ during exercise help identify ventilation-perfusion mismatch as well as hyperventilation.

@@

\subsection{Breathing frequency, $B_f$}
In healthy individuals, breathing frequency is one of the most important respiratory responses to exercise in order to increase oxygen uptake. However, a variety of respiratory disorders including chronic obstructive pulmonary disease or restrictive disorders such as interstitial lung disease can impair this response. Morbidly obese patients may also have additional restrictions on respiratory response due to their body habitus. On the other hand, hyperventilation in an otherwise healthy patient may make interpretation of the results difficult although the analysis of trends in other parameters such as $P_{ET_{O_2}}$ and $P{ET_{CO_2}}$ aids in the diagnosis of hyperventilation versus true respiratory problems.

@@

These findings have since been replicated in several other studies although the threshold value of $\dot{V}_{O_2}$AT as well as the inclusion of other CPET parameters to attribute risk differ in these studies. Hightower and co-workers studied 32 patients over the age of 18 undergoing a variety of major elective abdominal surgery. In this heterogeneous cohort, they reported that the percent predicted anaerobic threshold achieved $<$ 75\%, heart rate at the anaerobic threshold and the heart rate response from rest to anaerobic threshold were all independently associated with postoperative morbidity. While it is difficult to extrapolate results from this small heterogeneous cohort to the general surgical population, it would appear that other cardiopulmonary exercise test parameters may have a role in predicting risk in these patients.\parencite{hightower_pilot_2010}

Snowden and co-workers reported in a study of 171 patients of who 123 underwent surgery, that $\dot{V}_{O_2}AT<10.1$ ml/kg/min was associated with not only cardiovascular complications but also with other complications including pulmonary, renal, gastrointestinal, infective and haematological complications. One of the strengths of this study was the fact that the clinicians treating these patients were blinded to the results of preoperative cardiopulmonary exercise testing results thus avoiding management bias. They also included other cardiopulmonary exercise test parameters such as $\dot{V}_{O_2}$Peak and $\dot{V}_E/\dot{V}_{CO_2}$ at the anaerobic threshold in addition to $\dot{V}_{O_2}$AT in their analysis and showed that $\dot{V}_{O_2}$AT was more predictive of postoperative morbidity than POSSUM derived morbidity, cardiac risk scoring index or a validated activity questionnaire. However, this study also included a heterogeneous cohort of diseases including vascular, pancreatic and hepatobiliary disorders and sarcomas.\parencite{snowden_submaximal_2010}

@@

\subsection{Systemic inflammation and long-term survival}
Systemic inflammation is associated with poorer survival in patients undergoing potentially curative surgery for pancreatic cancer \parencite{jamieson_systemic_2005,clark_preoperative_2007,bhatti_preoperative_2010} as well as in patients with inoperable disease.\parencite{glen_evaluation_2006} Patients with ductal adenocarcinoma of the head of the pancreas undergoing potentially curative resection survived for a median of 21.5 months if their CRP was $\leq$ 10 mg/dl a month after their surgery but only 8.4 months if their CRP remained persistently elevated over 10 mg/dl approximately a month after their operation.\parencite{jamieson_systemic_2005} Similar findings have been reported in cancers involving other organs using both the mGPS and other scores such as the neutrophil-lymphocyte ratio (NLR). A selection of these studies are presented in Table %Create new table here for this.
\todo{Table for systemic inflammation and long-term survival}

@@

Preoperative systemic inflammation has been reported to be associated with infectious complications in patients undergoing potentially curative surgery for colorectal cancer.\parencite{moyes_preoperative_2009} In a study of 455 patients, Moyes and co-workers reported that an elevated preoperative modified Glasgow Prognostic Score (\ref{table:mGPS}) was associated with increased incidence of infectious complications in patients undergoing elective as well emergency colorectal cancer surgery. They postulated that several mechanisms may have a role including disregulation of cell-mediated immunity, impaired T-lymphocyte response, disorders in the complement pathway and possibly due to loss of lean tissue and protein as a consequence of systemic inflammation. Preoperative mGPS has also been shown to predict postoperative morbidity in patients undergoing oesophageal resection for cancer.\parencite{vashist_glasgow_2010}

@@

In a study of 383 patients undergoing elective rectal cancer surgery with primary anastomosis, Welsch and co-workers reported that persistently raised CRP level over 140 mg/L after the third/fourth postoperative day was associated with anastomotic leak.\parencite{welsch_c-reactive_2007} They also reported in a separate study of 688 patients undergoing pancreatic resection with pancreatico-jejunostomy for neoplastic disease or chronic pancreatitis, that persistently elevated CRP levels greater than 140 mg/L on the fourth postoperative day was associated with increased incidence of complications.

@@

While CRP level between the third and fifth postoperative day has been reported to be most predictive of complications,  the clinical signs do not become apparent until later in the postoperative course, often after the fifth day. This has led some authors to postulate that the elevated CRP levels may in fact be due to an abnormally modulated postoperative inflammatory response resulting in an initial exaggerated systemic inflammatory response syndrome (SIRS) followed by a compensatory anti-inflammatory response syndrome (CARS).

@@

This form of 'immunoparalysis' was first described in patients after major trauma with tissue damage \parencite{abraham_effects_1985,bandyopadhyay_negative_2007} or after haemorrhage on its own without associated tissue trauma. \parencite{stephan_hemorrhage_1987} In a detailed review of the mechanisms underlying the compensatory anti-inflammatory response syndrome, Ward and co-workers describe SIRS and CARS to be mirror images suggesting that a disproportionately high SIRS is followed by a period of immunosuppression that leaves the patient prone to further complications.\parencite{ward_compensatory_2008}

@@

Yamaguchi and co-workers compared the levels of pro- and anti-inflammatory cytokines in patients undergoing cholecystectomy versus patients undergoing trans-thoracic oesophagectomy. They reported that the initial inflammatory phase was followed by an immunosuppressive phase that started around the seventh postoperative day in patients undergoing oesophagectomy. However, patient who underwent an open cholecystectomy did not experience this immunosuppressive phase, leading them to postulate that the degree of immunosuppression was directly proportional to the initial pro-inflammatory process. This in turn was related to the greater degree of surgical stress and tissue trauma that occurs with a trans-thoracic oesophagectomy. They also reported that in a randomised cohort that received an infusion of lymphokine-activated natural killer cells immediately after oesophagectomy, there was a trend towards fewer infectious complications.\parencite{yamaguchi_postoperative_2006}

@@

The detrimental effects of jaundice on the heart and the circulatory system has been recognised for over 150 years now. King and co-workers reported that the intravenous injection of bile caused bradycardia, hypotension and eventually death in dogs.\parencite{king_effect_1909} The concept of a \textit{'jaundiced heart'} was fist put forth by Green and co-workers in 1986.\parencite{green_jaundiced_1986} They performed choledocho-caval anastomoses in 5 dogs and studied cardiac function before and 2 weeks after this procedure. They reported that 'cholemia' was associated with impaired left ventricular function and blunted response to sympathomimmetic agents. Similar findings have been reported by other authors in animal studies. \parencite{binah_obstructive_1985,bomzon_systemic_1986}

@@

\subsection{Role of preoperative biliary drainage}
\label{sec:preoperative_biliary_drainage}
%fang_pre-operative_2012 - Cochrane review
%wang_preoperative_2008 - Cochrane review
The increased incidence of adverse events after surgery in patients with obstructive jaundice resulted in routine preoperative biliary drainage becoming the standard practice in patients undergoing major pancreatic surgery. Moreover, while significant improvements in perioperative care has resulted in reduction in postoperative mortality, morbidity rates have remained high. Obstructive jaundice has provided a potential target for preoperative optimisation in an attempt to improve postoperative outcomes.

Early observational studies of small cohorts of patients challenged this paradigm reporting that surgery in jaundiced patients was not associated with greater risk and in fact preoperative biliary drainage increased hospitalisation costs. \parencite{snellen_influence_1985, bakkevold_morbidity_1993, pitt_does_1985} Hatfield and co-workers reported in one of the earliest randomised controlled trial comparing pancreaticoduodenectomy with or without preoperative biliary drainage in 57 patients that there was no difference in postoperative complications or mortality but increased biliary drainage associated complications. \parencite{hatfield_preoperative_1982}

@@

A recent randomised controlled trial of 202 patients reported that there was no significant difference in post-operative morbidity (47\% vs, 37\%, p=0.14), mortality or length of stay between the group that underwent preoperative biliary drainage and the group that underwent early surgery. There was a high incidence of procedure related complications in the drainage group (46\%). The high incidence of procedure related complications, the use of plastic stents, the 25\% failure to drain at first attempt and exclusion of patients with serum bilirubin $>$ 250 $\mu$mol/L were significant limitations of the study. However, the study does show that early surgery in the jaundiced patient with serum bilirubin $<$ 250 $\mu$mol/L is safe and not associated with increased morbidity or mortality as once feared. Mezhir and co-workers also reported a high (23\%) stent related complication rate in their retrospective study of 340 patients where 96\% of 201 patients in the biliary drainage group had plastic stents placed. However, they also reported a greater incidence of wound infections (7\% vs 20\%, p=0.01) and intra-abdominal abscesses (3\% vs 12\%, p=0.03) in patients who underwent biliary drainage.

@@

The impact of obesity on the rate of postoperative complications remains controversial. It has been a commonly held belief that obese patients had greater incidence of complications. Some of the adverse outcomes such as wound complications and respiratory dysfunction where considered to be the direct result of the increased body mass index while other complications where related to the associated comorbidities such as cardiovascular disease, obstructive sleep apnea or diabetes mellitus.
